---
title: "R Notebook"
output: html_notebook
---

Importing libraries
```{r}
library("tidyverse")
library("ggplot2")
library('readxl')
  
cpi = read_excel('CPIAUCSL.xls', skip = 10)
interest_rate = read_excel('FEDFUNDS.xls', skip = 10)

head(interest_rate)
```
```{r}
# To create the pct_change
cpi <- cpi %>%
  mutate(inflation_rate = (CPIAUCSL/lead(CPIAUCSL) - 1) * 100)

cpi
```



```{r}
data <- merge(cpi,interest_rate,by="observation_date")

head(data)
```
 
```{r}
ggplot(data,aes(FEDFUNDS, inflation_rate)) +
  geom_point() +
  geom_smooth(method='lm')
```

```{r}
library("writexl")
write_xlsx(data,"merged_data.xlsx")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

