---
title: "R Notebook"
output: html_notebook
---

# Observational Study of Historical Behavior of Interest and Inflation Rates in Brazil and USA

Push the changes to the github
OBS: $ git push -u origin main

IMPORTANT, BOTH CPI AND INTEREST RATE ARE ANNUALIZED FOR US

### Importing libraries
```{r}
library("tidyverse") #Manipulate Data
library("ggplot2") #Library for Data Visualization
library('readxl') #Read data from excel
library('GetBCBData') #Library to help to collect data from the BCB API (Central Bank of Brazil)
```
## Collecting USA Data

This excel files were downloaded from the FRED database of economical and financial data from USA.
[ADD DESCRIPTION OF THE DATA, SEARCH IF FROM THE FRED WEBSITE]
### Important notes:
+ The CPI is an index, to compute the annualized inflation rate, we need to compute (cpi_t12/cpi_0), since we have monthly data.
+ The interest rate is already annnualized
```{r}
# https://fred.stlouisfed.org/series/FEDFUNDS for description
cpi_us = read_excel('raw_data/CPIAUCSL.xls', skip = 10)
interest_rate_us = read_excel('raw_data/FEDFUNDS.xls', skip = 10)
cpi_us <- cpi_us %>%
  mutate(inflation_rate_annual =  CPIAUCSL/lag(CPIAUCSL, 12)-1 )# %>% LATER WHEN WE DECIDE TO COMPUTE THE CHANGE
  #mutate(change_inflation = (inflation_rate_annual-lag(inflation_rate_annual)))

inflation_rate_us <- cpi_us[c('observation_date', 'inflation_rate_annual')]

#Merge both databases
data_us <- merge(interest_rate_us, inflation_rate_us,by="observation_date")
colnames(data_us) <- c('observation_date', 'interest_us', 'inflation_us')

data_us$interest_us <- data_us$interest_us/100

head(data_us)
```

## Collecting Brazil Data
We will collect the CDI from Brazil as an proxy of the interest rate, and the IPCA which is a proxy of the Brazillian CPI. We will use a library to help us collect the data from the BCB API.
```{r}

#IPCA its brazillian CPI, it comes in monthly rate
brazil_cpi <- GetBCBData::gbcbd_get_series(
id = 433)
# TO ANNUALIZE THE DATA and save it in decimal format
brazil_cpi$value <- (brazil_cpi$value/100+1)^12-1 #To convert the data into annualized IPCA

# overnight rate in brazil annulized (base 252)
brazil_interest_rate <- GetBCBData::gbcbd_get_series(
id = 4389)
brazil_interest_rate$value <- brazil_interest_rate$value/100 #save it in decimal format

#merge both dataframes
data_br <- merge(brazil_cpi[c('ref.date','value')], brazil_interest_rate[c('ref.date','value')],by="ref.date")

colnames(data_br) <- c('observation_date', 'inflation_brazil', 'interest_brazil')

head(data_br)
```

## Descriptive Stats

```{r}
subset_us <- colnames(data_us)[-1]
subset_br <- colnames(data_br)[-1]

summary(data_us[subset_us])

summary(data_br[subset_br])

```
## Boxplot Data
```{r}
labels <- c(
  replicate(length(data_us$interest_us), 'Interest USA'),
  replicate(length(data_us$inflation_us), 'Inflation USA'),
  replicate(length(data_br$interest_br), 'Interest BR'),
  replicate(length(data_br$inflation_br), 'Inflation BR')
)

values <- c(
  data_us$interest_us, 
  data_us$inflation_us,
  data_br$interest_br,
  data_br$inflation_br
)

country <- c(
  replicate(length(data_us$interest_us)*2, 'USA'),
  replicate(length(data_br$interest_br)*2, 'Brazil')
)

boxplot_data <- data.frame(
  value =  values,
  category = labels,
  Country = country
)

head(boxplot_data)

ggplot(boxplot_data, aes(x=category, y=value, fill = Country)) +
  geom_boxplot() + 
  labs(
    y = 'Annualized Rate',
    x = 'Category',
    title = 'Destribution of Interest and Inflation Rates by Country'
  ) +
  scale_y_continuous(labels = scales::percent)
```

Save the data_frame in a xlsx file
```{r}
#library("writexl")
#write_xlsx(data,"merged_data.xlsx")
```

## Time Series for both Countries

Merge Data
```{r}

ggplot(data=data_br, aes(observation_date)) +
  #geom_line(aes(y = inflation_us, colour = "Inflation Rate USA"))+
  #geom_line(aes(y = interest_us, colour = "Interest Rate USA")) +
  geom_line(aes(y = interest_brazil, colour = "Interest Rate Brazil")) +
  geom_line(aes(y = inflation_brazil, colour = "Inflation Rate Brazil"))+
  ggtitle("Interest and Inflation Rates Over Time")+
  ylab("Annualized Rate") +
  xlab("Time") +
  scale_y_continuous(labels = scales::percent)
head(time_series)
```


```{r}
library(ggplot2)

ggplot(data=data_us, aes(observation_date)) +
  geom_line(aes(y = inflation_us, colour = "Inflation Rate USA"))+
  geom_line(aes(y = interest_us, colour = "Interest Rate USA")) +
  ggtitle("Interest and Inflation Rates Over Time")+
  ylab("Annualized Rate") +
  xlab("Time") +
  scale_y_continuous(labels = scales::percent)
head(time_series)
```

# Analyzing Relation between change in Interest Rate over Interest Rates
```{r}
# To create the inflation rate in pp%
  #cpi <- cpi %>%
   # mutate(inflation_rate_annual =  CPIAUCSL/lag(CPIAUCSL, 12)-1 )# %>% LATER WHEN WE DECIDE TO COMPUTE THE CHANGE
    #mutate(change_inflation = (inflation_rate_annual-lag(inflation_rate_annual)))
delta_us <- data_us %>% 
  mutate(change_inflation_us = inflation_us-lag(inflation_us, 1)) %>%
  mutate(change_interest_us = interest_us-lag(interest_us, 1))

delta_br <- data_br %>% 
  mutate(change_inflation_br = inflation_brazil-lag(inflation_brazil, 1)) %>%
  mutate(change_interest_br = interest_brazil-lag(interest_brazil, 1))

ggplot(data = delta_us, aes(x=change_interest_us, y=change_inflation_us)) +
  geom_hex()
```
```{r}
test <- data_us %>% 
  mutate(change_inflation_us = inflation_us/lag(inflation_us, 1)-1) %>%
  mutate(change_interest_us = interest_us/lag(interest_us, 1)-1)

ggplot(data = test, aes(x=change_interest_us, y=change_inflation_us)) +
  geom_point()
```


